\documentclass[landscape]{slides}

\usepackage[landscape]{geometry}
\usepackage{verbatim}
\usepackage{hyperref}

\author{James Jeffries}
\author{\parbox{.5\textwidth}{\centering
  James Jeffries \\
  \small Twitter: \href{http://www.twitter.com/jjeffries1}{@jjeffries1} \\[3pt]
  Github: \href{https://github.com/jajeffries/introToFSharp}{github.com/jajeffries/introToFSharp}}}

\title{A Brief Overview of F\#}
\date{7th June 2012}

\begin{document}

\begin{slide}
\maketitle
\end{slide}

\begin{slide}{Overview}
\begin{itemize}
\item Functional programming 
\item Features of F\#
\item Interoperating with .NET
\end{itemize}
\end{slide}

\begin{slide}{What is F\#?}
\begin{itemize}
\item A functional language base that runs on the .NET framework
\begin{itemize}
\item can call existing .NET libraries
\item can make libraries to be called by other .NET code
\end{itemize}
\item Shares the core of the language with OCaml
\end{itemize}
\end{slide}

\begin{slide}{What is functional programming and why use it?}
\begin{itemize}
\item Treats computation as the evaluation of functions and avoids state and mutable data
\item Avoiding state makes it easier to test and reason about code
\item If there are no dependencies between functions they can be performed in parallel
\item It's generally more conicse - writing less code makes it harder to write in bugs
\end{itemize}
\end{slide}

\begin{slide}{Why learn it?}
\begin{itemize}
\item Learning functional programming can improve you programming style when writing in other languages
\item It's fun!
\end{itemize}
\end{slide}

\begin{slide}{Hello World Explained}
\begin{verbatim}
let _ = println "Hello World"
\end{verbatim}
\begin{itemize}
\item let - binds an expression to a name
\item \_ - is the main or entry point
\item println - is the print function
\end{itemize}
\end{slide}

\begin{slide}{Functions}
\begin{verbatim}
let add a b = a + b
\end{verbatim}
\begin{itemize}
\item Type of this function is
\begin{verbatim}
val add : int -> int -> int
\end{verbatim}
\item Takes 2 ints and returns an int
\end{itemize}
\end{slide}

\begin{slide}{Currying}
\begin{verbatim}
let add a b = a + b
let addOne = add 1

let _ =
    printfn "%d" (addOne 3)
\end{verbatim}
\begin{itemize}
\item Currying allows us to effectively create a function with arguments already specified
\item addOne has the following type. It returns a function that takes an int and returns an int
\end{itemize}
\begin{verbatim}
val addOne : (int -> int)
\end{verbatim}
\end{slide}

\begin{slide}{Automatic generalisation}
\begin{itemize}
\item When the compiler does type inference it tries to determine which parameters can be generic and which need constraints.
\begin{verbatim}
> let max a b = if a > b then a else b ;;
val max : 'a -> 'a -> 'a when 'a : comparison
\end{verbatim}
\item Here the compiler has determined that the parameters can be generic, but must be able to be compared
\end{itemize}
\end{slide}

\begin{slide}{A More Detailed Example}
\begin{verbatim}
#light

let sqr x = x * x

let sumOfSquares nums =
    let mutable acc = 0
    for x in nums do
        acc <- acc + sqr x
    acc

let _ = sumOfSquares [1;2;3;4]
\end{verbatim}
\end{slide}

\begin{slide}{A More Detailed Example Explained}
\begin{description}
\item[\#light] use lightweight syntax
\item[mutable] declares a mutable variable. By default let binds an immutable value to a name.
\item[\textless-] puts the value of the expression on the right into the variable on the left
\item[acc] the last expression is the return a value
\end{description}
\end{slide}

\begin{slide}{A More Detailed Example Explained (continued...)}
\begin{itemize}
\item This is not good style as it is more imperative
\item Can be useful when first starting to write like this to get an understanding of the language
\item Also useful when using existing .NET libraries that aren't written in a functional style
\end{itemize}
\end{slide}

\begin{slide}{Taking a more functional approach}
\begin{verbatim}
#light

let sqr x = x * x

let rec sumOfSquares nums =
    match nums with
    | []   -> 0.0
    | h::t -> sqr h + sumOfSquares t

let _ = sumOfSquares [1;2;3;4]
\end{verbatim}
\end{slide}

\begin{slide}{Taking a more functional approach explained}
\begin{description}
\item[rec] declares a recursive function. Functions aren't recursive by default.
\item[match] matches a pattern. Sort of like a switch statement. match returns a value so we don't need to specify what to return seperately
\item[h::t] splits a list into it's head and tail. Binds the head to h and tail to t
\end{description}
\end{slide}

\begin{slide}{Using Pipelines}
\begin{verbatim}
#light

let sqr x = x * x

let sumOfSquares nums =
    nums 
    |> Seq.map sqr
    |> Seq.sum

let _ = sumOfSquares [1;2;3;4]
\end{verbatim}
\end{slide}

\begin{slide}{Using Pipelines Explained}
\begin{itemize}
\item pipelines are similar to pipes in unix
\item they allow you to pipe the result of the previous expression into the last argument of the next expression
\end{itemize}
\end{slide}

\begin{slide}{Using Lambdas}
\begin{verbatim}
#light

let sumOfSquares nums =
    nums 
    |> Seq.map (fun x -> x * x) 
    |> Seq.sum

let _ = sumOfSquares [1;2;3;4]
\end{verbatim}
\end{slide}

\begin{slide}{Using Lambdas Explained}
\begin{itemize}
\item lambdas are anonymous functions
\item declared using the fun keyword
\end{itemize}
\end{slide}

\begin{slide}{Lets make this parallel}
\begin{verbatim}
open System.Linq
open System

let pmap f (l:seq<_>) =
    let pe = l.AsParallel<_>()
    ParallelEnumerable.Select(pe, Func<_,_>(f))

let sumOfSquares nums =
    nums
    |> pmap (fun x -> x * x) 
    |> Seq.sum

let _ = sumOfSquares [1;2;3;4]
\end{verbatim}
\end{slide}

\begin{slide}{Lets make this parallel explained}
\begin{itemize}
\item As the functions don't rely on each other it is easy to parallelise using LINQ
\end{itemize}
\begin{description}
\item[open] opens a namespace. Similar to the using keyword in C\#
\item[l:seq\textless\_\textgreater] specifies the type of l
\item[\_] seq is generic but as we don't care what the type is we can use \_ as a wilcard
\end{description}
\end{slide}

\begin{slide}{Lets make this parallel explained (continued...)}
\begin{description}
\item[l.AsParallel\textless\_\textgreater()] calls seq's AsParallel method to get a ParallelEnumerable
\item[ParallelEnumerable.Select(pe, Func\textless\_,\_\textgreater(f))] Projects in parallel each element of a sequence into a new form
\end{description}
\end{slide}

\begin{slide}{Where can I learn more?}
\begin{itemize}
\item Expert F\# by Don Syme, Adam Granicz and Antonio Cisternino - very detailed and lots of good explanations
\item \href{http://channel9.msdn.com/blogs/pdc2008/tl11}{http://channel9.msdn.com/blogs/pdc2008/tl11} by Luca Bolognese - good overview to get you started with lots of useful tips
\item \href{http://msdn.microsoft.com/en-us/vstudio/hh386303}{http://msdn.microsoft.com/en-us/vstudio/hh386303} - lots of good videos and tutorials
\end{itemize}
\end{slide}

\end{document}
